---
title: 'rOCAT'
output: 
  github_document:
    toc: true
    number_sections: true
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

<!-- badges: start -->

[![Travis build status](https://travis-ci.com/bowang-lab/rOCAT.svg?branch=main)](https://travis-ci.com/bowang-lab/rOCAT)

<!-- badges: end -->

This is the R package for One Cell At A Time(OCAT), which provides a fast and memory-efficient framework for analyzing and integrating large-scale scRNA-seq data. Details of the methods can be check [here](https://genomebiology.biomedcentral.com/articles/10.1186/s13059-022-02659-1).

# Installation

You can install the development version of rOCAT from [GitHub](https://github.com/bowang-lab/rOCAT) with:

``` r
# install.packages("devtools")
devtools::install_github("bowang-lab/rOCAT")
```

# Example

Here we demonstrate how to sparsely encodes single-cell gene expression data on both single and multiple datasets

The **single** dataset using 3,005 cells and 4,412 genes in the mouse somatosensory cortex and hippocampal CA1 region from Zeisel et al. (2015) and can be download [here]().

The **multiple** datasets consist of five scRNA-seq datasets (Baron et al. 2016, Muraro et al. 2016, Segerstolpe et al. 2016, Wang et al. 2016, Xin et al. 2016) and can be download [here]().

## Working with Single Dataset

### load package

```{r example}
library(rOCAT)
library(reticulate)
```

### import data

```{r import}
# load Test Zeisel data
data <- R.matlab::readMat('../data/Test_5_Zeisel.mat')
# extract the gene feature matrix and make it saved as a sparse matrix
in_X <- as(data$in.X, 'dgCMatrix')
# extract the labels
labels_true <- as.vector(data$true.labs)

# the input data should be a vector of datasets c(datasets1,dataset2,...)
data_list <- c(in_X)
```

### run OCAT pipeline

```{r run, warning=FALSE, message=FALSE}
ZW <- run_OCAT(data_list, m_list=list(50), dim=30, 
                p=0.3, log_norm=TRUE, l2_norm=TRUE, if_inference=FALSE, 
                random_seed=42)

# first 5x5 dimension of ZW
ZW[1:5,1:5]
```

### clustering and evaluation

```{r run_OCAT}
labels_pred <- evaluate_clusters(np_array(ZW),num_cluster = length(unique(labels_true)))
normalized_mutual_info_score(labels_true, labels_pred)
```

## Working with Multiple Datasets

### import datasets

```{r import_multiple_dataset}
# the data is saved in .npz file in this example, so we need scipy from python to help read the data
scipy <- reticulate::import("scipy.sparse")
data_path <- "../data/Pancreas/data/"
files <- c('baron_1','muraro_2', 'seg_3', 'wang_4', 'xin_5')
data_list <- lapply(files,  FUN = function(x) scipy$load_npz(paste0(data_path,x,".npz")))

# import the labels
np <- reticulate::import("numpy") 
label_path <- "../data/Pancreas/label/"
labels <- lapply(files,  FUN = function(x) np$load(paste0(label_path,x,"_label.npy"),allow_pickle = TRUE))
labels_true_combined <- do.call(c, labels)
batch_true_combined <- rep(files,lengths(labels)) #batch labels 
```

### run OCAT pipline

```{r run_multiple_dataset, warning=FALSE, message=FALSE}
m_list <- list(65, 65, 65, 65, 65)
pancreatic_ZW <- run_OCAT(data_list, m_list, dim=60, p=0.3, log_norm=TRUE, l2_norm=TRUE)
pancreatic_ZW[0:5,0:5]
```

### clustering and evaluation

#### cell type clustering

```{r cell_type_cluster}
cell_type_pred <- evaluate_clusters(pancreatic_ZW, num_cluster = length(unique(labels_true_combined)))
normalized_mutual_info_score(labels_true_combined,cell_type_pred)
```

#### batch clustering

```{r batch_cluster}
batch_pred <- evaluate_clusters(pancreatic_ZW, num_cluster = length(files))
normalized_mutual_info_score(batch_true_combined,batch_pred)
```
