---
title: "trajectory-and-pseudotime-inference"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{trajectory-and-pseudotime-inference}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

In this page, we demonstrated how to run trajectory and pseudotime inference in [HSMM dataset](https://data.wanglab.ml/OCAT/HSMM.zip) with our rOCAT package

# load data and apply OCAT pipeline

## load package

```{r setup}
library(rOCAT)
library(reticulate)
library(Matrix)
```

## load data

```{r load data}
# load Test Zeisel data
data <- read.table('../../data/HSMM/HSMM.txt',header = TRUE,sep=' ',row.names = NULL)
row.names <- data$row.names
data <- data[,-1]
data <- Matrix::Matrix(as.matrix(data))
data <- t(data) # format as cell x Gene

data_list <- c(data)

ZW <- run_OCAT(data_list, m_list=list(40), dim=80)
```

# Trajectory inference

Load in the annotated labels.

```{r load_label}
#load labels
library(plyr)
labels <- data.frame(read.table('../../data/HSMM/HSMM_label.txt',row.names = NULL))
map <- data.frame(cbind(V1=c(1,2,3,4,5), 
          to=c("Fibroblast","Myotubes","Myoblasts","Undiff","Intermediates")))
labels_mapped <- plyr::join(labels,map,by='V1')$to
```

OCAT.compute_lineage() function infers Lineages over clusters with the OCAT features, predicted/true cluster labels and a user-specified root_cluster.

```{r Trajectory}
L <- compute_lineage(ZW,labels_mapped, root_cluster='Myoblasts', name='OE', reverse=0)
Lineage <- L[[1]]
root_cluster <- L[[2]]
cluster_labels <- L[[3]]
tree <- L[[4]]
```

# Pseudotime inference

OCAT.compute_ptime() function infers pseudotime for individual cell using the OCAT extracted features and the predicted lineage.

```{r Pseudotime}
P  <- compute_ptime(ZW, labels_mapped, Lineage, root_cluster)
Ptime <- P[[1]]
root_cell_list <- P[[2]]
```
